untitled

### 2016-05-11 model.tex
sample results
notes and code update

\subsection{Locally isothermal equation of state for the gas} 
We adopt 
\begin{align}\label{eos}
  P = c_s^2(r,z)\rhog = c_s^{2}(1 - \epsilon)\rho,   
\end{align}
where $c_s$ is a prescribed sound-speed profile fixed in time.

\subsection{Effective mixture energy equation}
We show that for a prescribed temperature distribution, the mixture
obeys an evolutionary energy equation, due to the advection of the
dust-fraction.

We eliminate $\epsilon$ in favor of $P$, 
\begin{align}
  \epsilon = 1 - \frac{P}{c_s^2(r,z)\rho}, 
\end{align}
then Eq. \ref{dusteq} becomes

\begin{align}
\frac{DP}{Dt} &= - P \nabla\cdot\bm{v} + P\bm{v}\cdot\nabla\ln{c_s^2}
                + \mathcal{C},  \label{eff_energy} \\
\mathcal{C}&\equiv c_s^2 \nabla\cdot\left[\tstop\left(1 - \frac{P}{c_s^2\rho}\right)\nabla
  P\right].
\end{align}
Eq. \ref{eff_energy} is the energy equation for an ideal gas of adiabatic index
$\gamma=1$, i.e. isothermal evolution, but now with additional source 
terms due to the imposed temperature profile, as well as dust-gas
friction.

>>>>

\subsection{Locally isothermal equation of state for the gas} 
We adopt 
\begin{align}\label{eos}
  P = c_s^2(r,z)\rhog = c_s^{2}(1 - \epsilon)\rho,   
\end{align}
where $c_s$ is a prescribed sound-speed profile fixed in time.  
%For analytic discussion we consider arbitrary functions of
%$c_s(r,z)$. 
In numerical calculations we will consider vertically  
isothermal disks with 
\begin{align}
  c_s^2(r) \propto r^{q},
\end{align}
where $q$ is the power-law index for the disk temperature.

\subsection{Effective energy equation}
We show that for a prescribed temperature distribution, the mixture
obeys an evolutionary energy equation, due to the advection of the
dust-fraction. Eliminate $\epsilon$ in favor of $P$, 
\begin{align*}
  \epsilon = 1 - \frac{P}{c_s^2(r,z)\rho}, 
\end{align*}
then Eq. \ref{dusteq} becomes
\begin{align}
\frac{DP}{Dt} &= - P \nabla\cdot\bm{v} + P\bm{v}\cdot\nabla\ln{c_s^2}
                + \mathcal{C},  \label{eff_energy} \\
\mathcal{C}&\equiv c_s^2 \nabla\cdot\left[\tstop\left(1 -
  \frac{P}{c_s^2\rho}\right)\nabla 
  P\right].
\end{align}
Eq. \ref{eff_energy} is the energy equation for an ideal gas of adiabatic index
$\gamma=1$, i.e. isothermal evolution, but now with additional source 
terms due to the imposed temperature profile, as well as dust-gas
friction.

### 2016-05-11 result.tex
notes and code update

\begin{figure}
\includegraphics[width=\linewidth]{figures/eigenvalues}
\caption{do not believe}
\end{figure}

>>>>

We consider a globally isothermal disk, $c_s=\mathrm{const.}$, by
setting $q=0$ in the equation of state (Eq. \ref{eos}), and a flat
mid-plane gas density, $\rhog(0) = \mathrm{const.}$, so there is no
global radial pressure gradient at the disk midplane. This setup
eliminates classic vertical shear instability and streaming
instabilities. The stopping time is
prescribed as 
\begin{align}
  \tstop = t_\mathrm{s0}\frac{\rho(z=0)}{\rho(z)}. 
\end{align}

As an example, we set the midplane dust-to-gas ratio 
$\tepsilon_0=0.02$, dust scale-height $H_\mathrm{d}=0.6\Hgas$, and
stopping time $t_\mathrm{s0}\OmK=0.02$. We consider modes with radial
wavenumber $k_x\Hdust=2\pi$. The vertical domain size is set to  
$\zmax=2\Hgas$. The aspect-ratio of the gas layer is
$h_\mathrm{g}=0.05$.

We use a pseudo-spectral method to solve  
Eq. \ref{lin_mass}---\ref{lin_energy} as a generalized eigenvalue
problem with $N=513$ grid points. Fig. \ref{eigen_map}   
shows a map of the eigenvalues in the $(\omega, s)$ plane. Note that
we restrict attention to modes with $|\sigma|\leq \OmK$. 
Eigenvalues on the `Y' branch should not be trusted. The 
high growth-rate modes are associated with disturbances concentrated
at the domain boundaries. Modes with $\omega\simeq 0$, i.e. purely
growing modes, are likely spurious because we generally expect
eigenvalues to be complex (see the integral relation in Appendix
\ref{var_prin}).

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues}
  \caption{Eigenvalues found for the fiducial isothermal dusty
    problem. 
  \label{eigen_map}} 
\end{figure}

We thus consider the slower-growing modes near the horizontal
axis. Fig. \ref{eigen_vec} show example eigenfunctions for a case with
frequency $(\omega,s)=(-2.79,0.20)\hgas\OmK$.

In Fig. \ref{eigen_2d}
we visualize this mode in the $(x,z)$ plane: colors show the
perturbation to the dust-fraction ($\dd\epsilon$), while vectors show
the perturbed meridional momenta. This mode has a growth
timescale of $\simeq 16$ orbits. This mode is unlikely driven by
vertical shear: the maximum vertical 
shear rate, $\mathrm{max}\left|r\p_z\Omega\right|\simeq10^{-2}\hgas\OmK$,
is much smaller than the mode growth rate $s\simeq 0.2\hgas\OmK$.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvec}
  \caption{Example eigenfunctions from the fiducial case. 
    \label{eigen_vec}} 
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{figures/result2d}
  \caption{Example eigenvectors from the fiducial case. The eigenvalue
    is $(\omega,s)=(-2.79,0.20)\hgas\OmK$.  
    \label{eigen_2d}} 
\end{figure}
